---
title: "Decision Trees 2"
output: html_notebook
---



```{r}
## Data Import

file_path <- "./data/spam.csv"
if (!file.exists(file_path)) {
  dir.create("./data")
  url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/spambase/spambase.data"
  download.file(url = url, 
                destfile = file_path)
}
```

```{r}
spam <- read.csv(file_path, sep = ",", header = F)
```

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(keras))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(rpart))
suppressPackageStartupMessages(library(rpart.plot))

source("./functions/train_val_test.R")
```
# Assign the column names
```{r}
col_names_to_set <- c("word_freq_make","word_freq_address","word_freq_all","word_freq_3d","word_freq_our","word_freq_over","word_freq_remove","word_freq_internet","word_freq_order","word_freq_mail","word_freq_receive","word_freq_will","word_freq_people","word_freq_report","word_freq_addresses","word_freq_free","word_freq_business","word_freq_email","word_freq_you","word_freq_credit","word_freq_your","word_freq_font","word_freq_000","word_freq_money","word_freq_hp","word_freq_hpl","word_freq_george","word_freq_650","word_freq_lab","word_freq_labs","word_freq_telnet","word_freq_857","word_freq_data","word_freq_415","word_freq_85","word_freq_technology","word_freq_1999","word_freq_parts","word_freq_pm","word_freq_direct","word_freq_cs","word_freq_meeting","word_freq_original","word_freq_project","word_freq_re","word_freq_edu","word_freq_table","word_freq_conference","char_freq_;","char_freq_(","char_freq_[","char_freq_!","char_freq_$","char_freq_#","capital_run_length_average","capital_run_length_longest","capital_run_length_total", "target" 
)
colnames(spam) <- col_names_to_set
```
# Check for missing values
```{r}
summary(spam)
```

```{r}
spam[is.na(spam), ]
```

```{r}
str(spam$target)
spam$target <- as.factor(spam$target)
```
## Train / Validation / Test Split
```{r}
set.seed(123)
c(train, val, test) %<-% train_val_test_split(df = spam, train_ratio = 0.8, val_ratio = 0.2, test_ratio = 0)
```

## Model Creation
```{r}
model_decision_tree <- rpart(formula = target ~ ., 
                             data = train)
```

## Visualisation
```{r}
rpart.plot::rpart.plot(model_decision_tree)
```

# Predictions

```{r}
train$target_pred <- predict(model_decision_tree, newdata = train)[, 2]
val$target_pred <- predict(model_decision_tree, newdata = val)[, 2]
```

```{r}
threshold <- 0.5
train$target_pred_class <- ifelse(train$target_pred >threshold, 1, 0) %>% as.factor()
val$target_pred_class <- ifelse(val$target_pred >threshold, 1, 0) %>% as.factor()
```

# Model Performance

## Baseline Classifier

```{r}
table(train$target)[1] / length(train$target) * 100

```
## Confusion Matrix

Calculate a confusion matrix for Training Data:

```{r}
threshold <- 0.5

train$target_pred <- predict(model_decision_tree, newdata = train)[, 2]
train$target_pred_class <- ifelse(train$target_pred >threshold, 1, 0) %>% as.factor()

conf_mat_train <- table(predicted = train$target_pred_class, actual = train$target)
conf_mat_train

```

Calculate a confusion matrix for Validation Data:

```{r}
conf_mat_val <- table(predicted = val$target_pred_class, actual = val$target)
conf_mat_val
```

Calculate the Accuracy from the confusion matrix (for training and validation data).

```{r}
caret::confusionMatrix(conf_mat_train)
caret::confusionMatrix(conf_mat_val)
```

