---
title: "Linear Regression: A Real Model"
output: html_notebook
---



```{r}
library(ggplot2)
library(dplyr)
library(corrr)
library(rsample)
library(recipes)
library(parsnip)
```

```{r}
diamonds
```
# Select all columns that are non-factors (numeric) and correlate
```{r}
diamonds %>%
  select(where(~ !is.factor(.x))) %>%
  corrr::correlate() %>%
corrr::fashion()
```
# Network Plot
```{r}
diamonds %>%
  select(where(~ !is.factor(.x))) %>%
  corrr::correlate() %>%
corrr::network_plot()
```
# Filtered variables with correlation greater than 50% 
```{r}
high_corr_varables <- diamonds %>%
  select(where(~ !is.factor(.x))) %>%
  correlate() %>%
  corrr::focus(price) %>%
  arrange(desc(price)) %>%
  filter(price > 0.5) %>%
  pull(term)
```
# If new column is added that has +50% correlation and notebook is run it will be added to high_corr_variables
```{r}
data <- diamonds %>%
  select(high_corr_varables, price)
data
```
# Above without correlation
```{r}
data <- diamonds
data
```


# Split the data
```{r}
data_split <- initial_split(data, strata = price)
data_train <- training(data_split)
data_test <- testing(data_split)
```
# Build model
```{r}
model <- lm(price ~ ., data = data_train)
```

# Non-Correlated results (more accurate in this case = more variables)
```{r}
summary(model)
```


# Correlated results
```{r}
summary(model)
```
# non-correlated Tidy format (categorical values have higher correlation than numeric values)
```{r}
broom::tidy(model) %>%
  arrange(p.value)
```


# correlated Tidy format
```{r}
broom::tidy(model)
```

# The Tidymodels Approach
# Split data
```{r}
data_split <- initial_split(data, strata = price)
data_train <- training(data_split)
data_test <- testing(data_split)
```

# Data Preprocessing (normalisation)
```{r}
data_rec <- data_train %>%
  recipe(price ~ ., data_train) %>%
  step_dummy(all_nominal()) %>%
  step_normalize(all_numeric(), -all_outcomes()) %>%
  prep()

juiced <- juice(data_rec)
```
# Build model
```{r}
lm_model <- parsnip::linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")
```
# Fit model
```{r}
lm_fit <- fit(lm_model, price ~ ., juiced)
lm_fit
```
# glance
```{r}
glance(lm_fit)
```
# Tidy
```{r}
tidy(lm_fit)
```
# Apply preprocessing on testing data
```{r}
results_train <- lm_fit %>%
  predict(new_data = juiced) %>%
  mutate(truth = data_train$price)
results_test <- lm_fit %>%
  predict(new_data = bake(data_rec, data_test)) %>%
  mutate(truth = data_test$price)

results <- results_train %>%
  mutate(type = "train") %>%
  bind_rows(
    results_test %>%
      mutate(type = "test")
  )
```
# Compare train and test accuracy
```{r}
results %>%
  group_by(type) %>%
  yardstick::rmse(truth, .pred)
```
# Visualise results
```{r}
ggplot(results, aes(truth, .pred)) +
  geom_point() +
  geom_abline(color = "red", size = 2) +
  coord_fixed()
```

