---
title: "Data Preprocessing"
output: html_notebook
---


```{r}
library(dplyr)
library(tidyr)
library(skimr)
```


```{r}
View(starwars)
skim(starwars)
```

```{r}
data <- starwars %>%
  select(height, mass, gender)
```
##Splitting
##Using rsample to automatically split the tsting and training data

```{r}
library(rsample)

data_split <- initial_split(data)
data_train <- training(data_split)
test_data <- testing(data_split)

```
##Feature Engineering
##Creating a BMI column from height and mass (mass divided by height squared)

```{r}
data_train <- data_train %>%
  mutate(BMI = mass / (height * height))
data_train
```
##Missing Values
##To see missing values
```{r}
skim(data_train)
any(is.na(data_train))
```
```{r}
colSums(is.na(data_train))
```
## If very few rows have missing values then they can be dropped. The rest can be filled with the mean if is.na
```{r}
data_tr_imputed <- data_train %>%
  drop_na(height, gender) %>%
  mutate(mass = ifelse(is.na(mass), mean(mass, na.rm = TRUE), mass),
  BMI = ifelse(is.na(BMI), mean(BMI, na.rm = TRUE), BMI))
skim(data_tr_imputed)
```
## Encoding Categorical Data 

```{r}
iris
skim(iris)
```
## Mutate factor of Species into an integer
```{r}
iris %>%
  mutate(Species = as.integer(Species))
```
## One-hot Encoding (if Species is versicolor then 1 else 0) & same with virginica
## Third species is automatically encoded
```{r}
iris %>%
  mutate(Species_versicolor = ifelse(Species == "versicolor", 1, 0),
         Species_virginica = ifelse(Species == "virginica", 1, 0)) %>%
  select(-Species)
```
## Encoding Gender in Starwars (if masculine 1 else if feminine 0)
```{r}
data_tr_imputed_encoded <- data_tr_imputed %>%
  mutate(gender_masculine = ifelse(gender == "masculine", 1, 0)) %>%
  select(-gender)
data_tr_imputed_encoded
```
## Feature Scaling
```{r}
normalize <- function(feature) {
  (feature - mean(feature)) /sd(feature)
}
```

## Apply normilisation as all values are numeric
```{r}
data_training_imputed_encoded_normalized <- data_tr_imputed_encoded %>%
  mutate_all(normalize)
data_training_imputed_encoded_normalized
```

## Complete Preprocessing Pipeline from all previous steps
```{r}
data_train_pipeline <- data_train %>%
  mutate(BMI = mass / (height * height)) %>%
    drop_na(height, gender) %>%
  mutate(mass = ifelse(is.na(mass), mean(mass, na.rm = TRUE), mass),
  BMI = ifelse(is.na(BMI), mean(BMI, na.rm = TRUE), BMI)) %>%
    mutate(gender_masculine = ifelse(gender == "masculine", 1, 0)) %>%
  select(-gender) %>%
  mutate_all(normalize)
data_train_pipeline
  
```
## Comparing Pipeine to previous steps to see they are the same
```{r}
waldo::compare(data_train_pipeline, data_training_imputed_encoded_normalized)
```
## Do all of the previous steps using recipies (much more efficient)

```{r}
library(recipes)

data_recipe <- data_train %>%
  recipe() %>%
  step_mutate(BMI = mass / (height * height)) %>%
  step_naomit(height, gender) %>%
  step_impute_mean(mass, BMI) %>%
  step_dummy(gender) %>%
  step_normalize(everything()) %>%
  prep()
```

##To view the recipe use juice()
```{r}
data_preprocessed <- juice(data_recipe)
```
## Compare recipe to results from manual steps
```{r}
waldo::compare(data_preprocessed, data_train_pipeline)
waldo::compare(data_preprocessed, data_training_imputed_encoded_normalized)
```

