---
title: "Week 5: Decision Trees and Random Forests"
output: html_notebook
---

# Import data

```{r}

train <- read.csv("train.csv", stringsAsFactors = FALSE)
test <- read.csv("test.csv")

```

# Exploring datasets

```{r}

str(train)

table(train$Survived)

prop.table(table(train$Survived))

test$Survived <- rep(0, 418) # create table in test


```

# Women and children first?

```{r}

train$Sex <- as.factor(train$Sex)

summary(train$Sex) # Majority of passangers were male

prop.table(table(train$Sex, train$Survived), 1) # How many of each gender survived?


```

# Update prediction

```{r}
test$Survived <- 0

test$Survived[test$Sex == 'female'] <- 1

```

# Age

```{r}

summary(train$Age)

train$Child <- 0 # Create new variable child
test$Child <- 0

train$Child[train$Age < 18] <- 1 # Add all people under 18 from age to child
test$Child[test$Age < 18] <- 1 # Add all people under 18 from age to child

aggregate(Survived ~ Child + Sex, data = train, FUN = function(x){sum(x) / length(x)})

```

# Fare

```{r}

train$Fare2 <- '30+'
test$Fare2 <- '30+'

train$Fare2[train$Fare < 30 & train$Fare >= 20] <- '20-30'
train$Fare2[train$Fare < 20 & train$Fare >= 10] <- '10-20'
train$Fare2[train$Fare < 10] <- '<10'

test$Fare2[test$Fare < 30 & test$Fare >= 20] <- '20-30'
test$Fare2[test$Fare < 20 & test$Fare >= 10] <- '10-20'
test$Fare2[test$Fare < 10] <- '<10'

aggregate(Survived ~ Fare2 + Pclass + Sex, data = train, FUN = function(x) {sum(x)/length(x)})

```

# New prediction

```{r}

test$Survived <- 0

test$Survived[test$Sex == 'female'] <- 1

test$Survived[test$Sex == 'female' & test$Pclass == 3 & test$Fare >= 20] <- 0

```

# Decision Trees

```{r}

library(rpart)

fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked,
             data = train,
             method = "class")

plot(fit) # Ugly!
text(fit)

library(rattle)
library(rpart.plot)
library(RColorBrewer)

fancyRpartPlot(fit) # Much better!

```

# Prediction

```{r}

Prediction <- predict(fit, test, type = "class")
confMat <- table(test$Survived, Prediction)
accuracy <- sum(diag(confMat)) / sum(confMat)
accuracy * 100

```

# Purposely Overfitting

```{r}

fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked,
             data = train,
             method = "class",
             control = rpart.control(minsplit = 2, cp = 0))

fancyRpartPlot(fit)

Prediction <- predict(fit, test, type = "class")
confMat <- table(test$Survived, Prediction)
accuracy <- sum(diag(confMat)) / sum(confMat)
accuracy * 100

```

# Control Parameters

```{r}

fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked,
             data = train,
             method = "class",
             control = rpart.control(minsplit = 24, cp = 0))

new.fit <- prp(fit, snip = TRUE)$obj

fancyRpartPlot(new.fit)
```

# Feature Engineering

```{r}

train$Name <- as.factor(train$Name)
train$Name[1]

# Merge train / test

test$Survived <- NA
combi <- rbind(train, test)

```

```{r}
# Change back to a char

combi$Name <- as.character(combi$Name)
combi$Name[1]

# Split the char apart
strsplit(combi$Name[1], split = '[,.]')

strsplit(combi$Name[1], split = '[,.]')[[1]]

strsplit(combi$Name[1], split = '[,.]')[[1]][2] # index 2

combi$Title <- sapply(combi$Name, FUN = function(x) {strsplit(x, split = '[,.]')[[1]][2]}) # Create Title variable (column)

combi$Title <- sub(' ','', combi$Title) # Strip spaces from beginning of titles

table(combi$Title)
```

# Redundant titles

```{r}

combi$Title[combi$Title %in% c('Mme', 'Mlle')] <- 'Mlle' # Combine lesser used titles

combi$Title[combi$Title %in% c('Capt', 'Don', 'Major', 'Sir')] <- 'Sir'
  
combi$Title[combi$Title %in% c('Dona',  'Lady', 'the Countess', 'Jonkheer')] <- 'Lady'

combi$Title <- factor(combi$Title) # Change back to a factor

```


# Combining family

```{r}

combi$FamilySize <- combi$SibSp + combi$Parch + 1

```


# Combining Surname and FamilySize

```{r}
# Extract passenger lastnames

combi$Surname <-sapply(combi$Name, FUN = function(x) {strsplit(x, split = '[,.]')[[1]][1]})

# Change FamilySize to a char and append 

combi$FamilyID <- paste(as.character(combi$FamilySize), combi$Surname, sep = "")

# Any family of size <2 are classed as small (pesky Johnsons)

combi$FamilyID[combi$FamilySize <= 2] <- 'small'

famIDs <- data.frame(table(combi$FamilyID))

```


# Show only families with 2 or less members

```{r}

famIDs <- famIDs[famIDs$Freq <=2, ]

```


# Overwrite fasmily IDs in combi that were not correctly identified and convert into a factor

```{r}

combi$FamilyID[combi$FamilyID %in% famIDs$Var1] <- 'small'
combi$FamilyID <- factor(combi$FamilyID)

```


# Train / Test split

```{r}

train <- combi[1:891,]
test <- combi[892:1309,]

```


# Fit model

```{r}

fit <- rpart(Survived ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked + Title + FamilySize + FamilyID,
             data = train,
             method = "class")

fancyRpartPlot(fit)

```


# Random Forest

```{r}

# Fill NA's in Age

summary(combi$Age) # 263 NA's

Agefit <- rpart(Age ~ Pclass + Sex + SibSp + Parch + Fare + Embarked + Title + FamilySize,
                data = combi[!is.na(combi$Age),],
                method = "anova")

combi$Age[is.na(combi$Age)] <- predict(Agefit,
                                       combi[is.na(combi$Age),]) 

summary(combi$Age)

```

# Fix Embarked

```{r}

summary(combi) # Combi is char and Fare has an NA

combi$Embarked <- as.factor(combi$Embarked) # Convert to a factor

summary(combi$Embarked) # Embarked has a blank

which(combi$Embarked == '') # Indexes of blank fields (62, 830)

combi$Embarked[c(62,830)] = "S" # Fill the blanks with 'S'
combi$Embarked <- factor(combi$Embarked)

```

# Fix Fare

```{r}

summary(combi$Fare)

which(is.na(combi$Fare)) # which index (1044)

combi$Fare[1044] <- median(combi$Fare, na.rm = TRUE)

```

# Reduce number of levels in FamilyID

```{r}

combi$FamilyID2 <- combi$FamilyID
combi$FamilyID2 <- as.character(combi$FamilyID2)
combi$FamilyID2[combi$FamilySize <= 3] <- 'small'
combi$FamilyID2 <- factor(combi$FamilyID2)

```

# Test / Train Split
```{r}
train <- combi[1:891,]
test <- combi[892:1309,]
```

# Fit the model

```{r}

library(randomForest)

set.seed(415)

fit <- randomForest(as.factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + Fare +
                    Embarked + Title + FamilySize + FamilyID2,
                    data = train,
                    importance = TRUE,
                    ntree = 2000)

varImpPlot(fit)

```

# Predictions

```{r}

Prediction <- predict(fit, test)


```

# Forest of conditional inference trees

```{r}

library(party)

set.seed(415)

fit <- cforest(as.factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch +
                 Fare + Embarked + Title + FamilySize + FamilyID,
               data = train,
               controls = cforest_unbiased(ntree = 2000, mtry = 3))


Prediction <- predict(fit, test, OOB = TRUE, type = "response")

```

