---
title: "Week 8: Support Vector Machine"
output: html_notebook
---

# Load packages

```{r}

library('caret')

# load the data set:

heart <- read.csv("heart_disease.csv", sep = ',', header = FALSE)

```

In the above line of code, we’re reading the dataset which is stored in a CSV format
and that’s why we’ve used the read.csv function to read it from the specified path
The ‘sep’ attribute indicates that the data is stored in a CSV or Comma Separated
Version

```{r}
# Check the structure of our dataset

str(heart)

```

Our next step is to split the data into training set and testing set, this is also called
data splicing.
We’ll be using the training set specifically for our model building and the testing set
for evaluating the model:


# Split data into test / train

```{r}

intrain <- createDataPartition(y = heart$V14, p= 0.7, list = FALSE)
training <- heart[intrain,]
testing <- heart[-intrain,]

```

The caret package provides a method createDataPartition() which is basically for partitioning
our data into train and test set.

We’ve passed 3 parameters to this createdatapartition() function:
 
The “y” parameter takes the value of variable according to which data needs
to be partitioned. In our case, target variable is at V14, so we are passing
heart$V14

The “p” parameter holds a decimal value in the range of 0-1. It’s to show the
percentage of the split. We are using p=0.7. It means that data split should be
done in 70:30 ratio. So, 70% of the data is used for training and the remaining
30% is for testing the model.

The “list” parameter is for whether to return a list or matrix. We are passing
FALSE for not returning a list.


Now this createDataPartition() method is returning a matrix “intrain”. This intrain matrix has
our training data set and we’re storing this in the ‘training’ variable and the rest of the data,
i.e. the remaining 30% of the data is stored in the testing variable.

Next, for checking the dimensions of our training data frame and testing data frame, we can
use these:

```{r}

dim(training)

dim(testing)

```

Our next step is to clean the data, so if there are any missing values or inconsistent
values, they have to be dealt with before we build the training model
We’ll be using the anyNA() method, which checks for any null values:

```{r}

anyNA(heart)

```

on running this, we get the return values as false, which means that there are no
missing values in our dataset.
Next, we’re checking the summary of our data by using the summary() function

```{r}

summary(heart)

```

The output shows that the values of the various variables are not standardized.
For example, the V14 variables, which is our target variable, it holds only 2 values,
either 0 or 1.
Instead, this should be a categorical variable. To convert these to categorical
variables, we need to factorize them:

# Change into categorical variables

```{r}

training[["V14"]] = factor(training[["V14"]])

```

Our next step is to train our model.

Before we train our model, we’ll first implement the trainControl() method. This will
control all the computational overheads so that we can use the train() function
provided by the caret package. The training method will train our data on different
algorithms.

traincontrol() method:

```{r}

trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
                       
```

The trainControl() method here, is taking 3 parameters.

The “method” parameter defines the resampling method, in this demo we’ll be
using the repeatedcv or the repeated cross-validation method.

The next parameter is the “number”, this basically holds the number of
resampling iterations.

The “repeats ” parameter contains the sets to compute for our repeated cross-
validation. We are using setting number =10 and repeats =3

This trainControl() method returns a list. We are going to pass this on our train()
method.

```{r}

svm_Linear <- train(V14 ~., data = training, method =
"svmLinear",trControl=trctrl,preProcess = c("center", "scale"),tuneLength
= 10)

```

The train() method should be passed with “method” parameter as “svmLinear”. We
are passing our target variable V14. The “V14~.” denotes a formula for using all
attributes in our classifier and V14 as the target variable. The “trControl” parameter
should be passed with results from our trianControl() method. The “preProcess”
parameter is for preprocessing our training data

We are passing 2 values in our “pre-process” parameter “center” & “scale”.

```{r}

svm_Linear

```

