---
title: "Working with data"
output: html_notebook
---

# Data Preparation

Before you can visualise your data, you have to get it into R. This involves importing the data from an external source and converting it into a useful format.

## Importing data

R can import data from almost any source, including text files, excel spreadsheets, statistical packages, and database management systems. We'll  illustrate these techniques using the Salaries dataset, containing the 9-month academic salaries of college professors at a single institution in 2008-2009.

# Text files

The readr package provides functions for importing delimited text files into R data frames.

```{r}
library(readr)
```

# import data from a comma delimited file

```{r}
Salaries <- read.csv("salaries.csv")
```

# import data from a tab delimited file

```{r}
Salaries <- read.table("salaries.txt")
```

These functions assume that the first line of data contains the variable names, values and are separated by commas or tabs respectively, and that missing data are represented by blanks.

The readxl package can import data from Excel workbooks. Both xls and xlsx formats are supported.

```{r}
library(readxl)
```

# import data from an Excel workbook

```{r}
Salaries <- read_excel("salaries.xlsx", sheet=1)
```

